---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# astr

<!-- badges: start -->
<!-- badges: end -->

The goal of astr is to make working with Access Sensor Technologies (AST) air sampler log files easy.

## Installation

The `astr` package can be installed from our [GitHub](https://github.com/accsensors/astr).
You must have the `devtools` package installed.
Use the following commands to install the `devtools` and `astr` packages.

```r
install.packages("devtools")
devtools::install_github("accsensors/astr")
```

## Reading Air Sampler Log Files into R

Once the `astr` package has been installed, you can load the package using the following command:

```{r load_pkg}
library(astr)
```

Two functions are used to read log files into R: `read_ast_header()` and `read_ast_log()`.  These functions will read UPAS v2, UPAS v2.1, UPAS v2.1 PLUS, and Home Health Box (HHB) v2 log files.  Use the following commands to view the documentation for these two functions:  

```{r documentation}
?read_ast_header
?read_ast_log
```

You can use either the `base::lapply()` function or the `purrr::map()` function to read header data from multiple log files and combine those data into a single data frame with one row for each file. Below is an example using `lapply()`. If you prefer to use `purrr::map()`, just replace "lapply" with "purrr::map" in the example below. 

```{r header_lapply}
data_upas_h <- list.files(path = ".", pattern = "^PS.*.txt$", full.names = T) %>%
  lapply(read_ast_header, update_names = T)%>%
  dplyr::bind_rows()
```

You can also use either the `base::lapply()` function or the `purrr::map()` function to read sample logs from multiple files and combine those data into a single data frame. A column with the file name will be appended so that the data from each individual file can be identified easily.  

```{r log_lapply}
data_upas <- list.files(path = ".", pattern = "^PS.*.txt$", full.names = T) %>%
  lapply(read_ast_log, update_names = T) %>%
  dplyr::bind_rows()
```

For the examples above, you will need to update the `path` argument in the `list.files()` function to indicate the location where your log files are stored, relative to your working directory. Specifying `pattern = "^PS.*.txt$"` will choose only UPAS log files, since UPAS v2, UPAS v2.1, and UPAS v2.1 PLUS log files are prefixed by "PS", "PSP", and "PSP" respectively. To read only UPAS v2, UPAS v2.1 and UPAS v2.1 PLUS, or HHB v2 files, edit the `pattern` argument in `list.files()` as shown below:  

```{r specific_files}
# UPAS v2 files
data_upasv2 <- list.files(path = ".", pattern = "^PS[1-9].*.txt$", full.names = T) %>%
  lapply(read_ast_log, update_names = F) %>%
  dplyr::bind_rows()

# UPAS v2.1 and UPAS v2.1 PLUS files
data_upasv2x <- list.files(path = ".", pattern = "^PSP.*.txt$", full.names = T) %>%
  lapply(read_ast_log, update_names = F) %>%
  dplyr::bind_rows()

# HHB v2 files
data_hhb <- list.files(path = ".", pattern = "^HHB.*.csv$", full.names = T) %>%
  lapply(read_ast_log, update_names = T)%>%
  dplyr::bind_rows()
```

This package also includes many example log files to test your scripts. To see a list of the example log files included with this package, use the command below:

```{r example_files}
example_files <- system.file("extdata", package = "astr", mustWork = T) |>
    list.files(pattern = NULL, full.names = T)
```

You will notice that example log files from UPAS v2 ("PS" prefix), UPAS v2.1 PLUS ("PSP" prefix), and HHB v2 ("HHB" prefix) are included. 

Use the commands below to generate example data frames using the example log files. Change the `pattern` argument in `list.files()` as previously described to select the types of log file read (e.g., `pattern = "^HHB.*.csv$"` for HHB log files).

```{r example_header}
data_upas_examples_h <- system.file("extdata", package = "astr", mustWork = T) |>
    list.files(pattern = "^PS.*.txt$", full.names = T) %>%
    lapply(read_ast_header, update_names = T) %>%
    dplyr::bind_rows()
```

```{r example_log}
data_upas_examples <- system.file("extdata", package = "astr", mustWork = T) |>
    list.files(pattern="^PS.*.txt$", full.names = T) %>%
    lapply(read_ast_log, update_names = T) %>%
    dplyr::bind_rows()
```

## Make a GPS Map of UPAS Data

The `astr` package includes the experimental function `gps_map()` for making a GPS map of PM~2.5~ or CO~2~ data. This function is mainly for use with our online [Shiny App](https://accsensors.shinyapps.io/shinyAST/), but can be used in general scripts. Since `gps_map()` is experimental, the appearance and functionality of the map may change during future development. 

```{r example_map}
multiple_upas_logs <- system.file("extdata", package = "astr", mustWork = TRUE) |>
    list.files(pattern = "^PS.*.txt$", full.names = TRUE) %>%
    lapply(read_ast_log, update_names=TRUE) %>%
    dplyr::bind_rows()

upasv2x_pm25_map <- gps_map(multiple_upas_logs, variable = "PM2_5MC")
upasv2x_CO2_map  <- gps_map(multiple_upas_logs, variable = "CO2")
```

## Other Helper Functions

The `astr` package contains helper functions called by the `read_ast_header()`, and
`read_ast_log()` functions to process files. If you would like to use any of these helper functions independently in your script, you can find documentation and examples by typing a '?' before the function name.
These helper functions are: `count_header_rows()`, `format_ast_header()`, `format_ast_log()`, `format_hhb_header()`, `format_hhb_log()`, `format_upasv2_header()`, `format_upasv2_log()`, `format_upasv2x_header()`, `format_upasv2x_log()`, `fread_ast_header()`, `fread_ast_log()`, and `transpose_ast_header()`. 

The `astr` package also includes functions we use for our online [Shiny App](https://accsensors.shinyapps.io/shinyAST/). We do not recommend using these functions in your scripts because they are experimental and likely to change during future development.  These functions are: `shiny_axis()`, `shiny_header()`, `shiny_log()`, `shiny_sample_operation()`, `shiny_sample_settings()`, `shiny_sample_summary()`, `shiny_success_flag()`, and `shiny_units()`. 
